---
title: "Chapter 6 Lab 5 - Archetypal Analysis (AA)"
output:
  html_document:
    df_print: paged

---

# Goal
In this lab, we will show you how to use `Archetypal Analysis (AA)` to obtain archetypes from the data.

We will use the AA implementation from the "archetypes" package. See url: https://cran.r-project.org/web/packages/archetypes/archetypes.pdf for more information on the package. 

## Preparation

Load the following packages (install if you need to). This is similar to the previous labs, except we are now loading and will be using archetypes package. 

```{r}

options(repos = c(CRAN = "http://cran.rstudio.com"))
install.packages("archetypes", dependencies = TRUE)
library(archetypes) # for archetypal analysis
library(caret) # for preprocessing the data
```

# Step 1: Load the data set 

Similar to previous labs, load data and convert to appropriate data type.

```{r}

# data
Dota_Data <- read.csv("DoTalicious_cleaned1000players.csv", fileEncoding="UTF-8-BOM")
summary(Dota_Data)

# convert TotalTime
Dota_Data$TotalTime <- as.numeric(Dota_Data$TotalTime)

# exclude the row with SillLevel = SkillLevelNull
Dota_Data$SkillLevel <- as.factor(Dota_Data$SkillLevel)
Dota_Data <- Dota_Data[!(Dota_Data$SkillLevel == " SkillLevelNull"),]

# use factor in place of as.factor to remove " SkillLevelNull" from the list of factors
Dota_Data$SkillLevel <- factor(Dota_Data$SkillLevel)

# exclude PlayerID
Dota_Data <- Dota_Data[-1]

```

Similar to previous labs, assign the `SkillLevel` to `row.names` for visualization later.

```{r}
# make the names of levels descriptive and assign them to row.names
levels(Dota_Data$SkillLevel) <- c("N", "B", "I", "A")
row.names(Dota_Data) <- make.names(Dota_Data[,"SkillLevel"], unique = TRUE) 

```

Similar to previous labs, exclude the `SkillLevel` column, since it is not numeric. Archetypcal clustering algorithm only works with numeric variables.

```{r}

attribs_except_skill <- -which(names(Dota_Data) %in% c("SkillLevel"))
DotaData_NoSkill <- Dota_Data[, attribs_except_skill]

```

Finally, similar to previous labs, scale the data. 
```{r}
preProcValues <- preProcess(x = DotaData_NoSkill,method = c("center", "scale"))
DotaData_NoSkill <- predict(preProcValues, DotaData_NoSkill)
```

Check if there are any missing values (NA here means Not Available). If it returns *False*, we're good to go. We will leave out the rows with NAs, if they exist. 
```{r}
anyNA(DotaData_NoSkill)
DotaData_NoSkill = na.omit(DotaData_NoSkill)
```

# Step 2: Apply the Archetypal Analysis

We use the `stepArchetypes` function from the `archetypes` library. This function runs AA multiple times with different values of `K`, the number of archetypes.

```{r}
set.seed(2063)

# run archetypal analysis multiple times with different K, in this line we are considering values of 1-10 for K. 
ArchetypesForDota <- stepArchetypes(data = DotaData_NoSkill, k = 1:10, verbose = FALSE, nrep = 4)
```

We tried 10 different values of `K`, from 1 to 10, and repeated the `AA` algorithm 4 times per `K`. Since solving for archetypes is stochastic, it can converge to suboptimal solutions, thus repeating it multiple times of `K` helps ensure that we do not get stuck in local optima.

If you see some warnings, those indicate that the matrices were singular, which can be caused by some unfavorable random initializations, and can be ignored.

Next, the `screeplot` function shows the plot of RSS (Residual Sum of Squares) scores w.r.t different numbers of archetypes. The lower the RSS score is, the better the model is.

```{r}
screeplot(ArchetypesForDota)
```

In this case, 9 can be the optimal number of archetypes. However, you would also want to set a small number of archetypes. As such, the Elbow method can be applied here, suggesting `K=6` as a reasonable number of archetypes as well. 

To obtain the resultant archetypes with respect to different `K` values, we use the `bestModel` function. For example, the following shows information related to the model at `K=10`.

```{r}

num_archs <- 10
BestArchetypeModelForDota <- bestModel(ArchetypesForDota[[num_archs]])
BestArchetypeModelForDota
```

And this shows that the Archetypes obtained with `K=10`.

```{r}
BestArchetypeModelForDota$archetypes
```

Let's take a look at the data by raw and their alpha coefficients given Archetype. This will help us establish which Archetype each data point belongs to.

```{r}
BestArchetypeModelForDota$alphas
```


# Step 3: Visualization

Archetypes are different from clusters in the traditional sense, and more similar to principle components in PCA.

One way to visualize archetypes is by using `simplexplot`.

```{r}

simplexplot(bestModel(ArchetypesForDota[[num_archs]]))

```

In the plot, the archetypes are shown on the perimeter of a simplex, named as `A_1` to `A_K` (K=10 in this case), containing the data points. The distances between each data point to the archetypes represent the "closeness" it is to the archetypes.

Let's try with a smaller K, such as 6.

```{r}

simplexplot(bestModel(ArchetypesForDota[[6]]))
```


---
**REMARKS**

Similar to the case with Model-based clustering, AA cannot be compared with other traditional clustering algorithms, since its goal, thus the objective function it is optimizing, is different. 

---


# Conclusion
In this lab, we showed you how to use Archetypal analysis to extract archetypes from a data set. This type of algorithm is very different from others, so you can compare the results qualitatively to other algorithms and see what results in better qualitative clusters.  
