---
title: "Chapter4_Lab2"
output:
  html_document:
    df_print: paged
---
# Goal

Is to replicat the PCA lab (Lab 1) but with both categorical and numerical variables. For demonstration we will use the Dota data we used in the previous lab, with two categorical variables: `skill level` and `gender`. It should be noted that while `skill level` is in the original dataset, we added the gender to this dataset so we can demonstrate what you can do when you have more than one categorical variable. `Skill level` is a factor variable with values: `-1`, `0`, `1`, or `2`. Although it is represented numerically, it is a categorical variable as the distance between 0 and 1 or -1 is not consistent. 

Let's first read in the table. 

```{r}
#We are going to be using the PCA library so we need to load that up
library(PCAmixdata)

#read the table and display the head of the table
Dota_datamixed <- read.table("Dota_normalized_mixed.csv", header=TRUE, sep=",")
head(Dota_datamixed)
```

#Grouping Numerical and Categorical 

We then separate the data into categorical variables and numerical variables. 

```{r}
#two sets numerical and categorical
Dota_Numerical <- Dota_datamixed[,1:15]
Dota_Numerical[1:10,]

Dota_Categorical <- Dota_datamixed[,16:17]
Dota_Categorical$SkillLevel <- as.factor(Dota_Categorical$SkillLevel)
Dota_Categorical[1:10,]
```

#Run PCAmix

We first run the PCAmix and then look at the Eigen vectors on the dimensions. 
Once you see that, you can see that 5-6 dimensions explains most of the variance, with 90% at dim =7 and 86% at dim=6. So we chose 6 dimensions. 

Similar with PCA, we can also get the variables in the new coordinates as shown in the code below. 
```{r}
#PCAmix
Dota_PCAmix <- PCAmix(X.quanti = Dota_Numerical, X.quali = Dota_Categorical, ndim=6, graph=FALSE)

#eigen vectors and their variance
Dota_PCAmix $eig

#variables in the new coordinates
head(Dota_PCAmix$ind$coord )

```

#Plotting

This plot will plot the link between the variables and the principle components. As you can see, almost all numerical variables except `KillsPerMin` and `Points` are linked to the first component (shown as Dim 1 in the graph). But `KillsPerMin` is almost orthogonal to the first component. `Skill Leve`l and `Points` are both linked to both second Principle Component and the first one. We can further inspect the loadings to understand the values per variable for each new component. 
```{r}
#plotting all variables
plot(Dota_PCAmix ,choice="sqload",coloring.var=T, leg=TRUE, posleg="topright", main="(d) All variables")
```

#Loadings 

We can also get the loadings for the quantitative and qualitative coordinates. 
```{r}

Loadings_quant <- Dota_PCAmix $quanti.cor
Loadings_quant

Loadings_qual <- Dota_PCAmix $categ.coord
Loadings_qual

```

In conclusion, through this lab, you have seen how to apply PCA on a mixed categorical and numerical data types. You can then use the PCs as you did in Lab 1. 