---
title: "Chapter 7 Lab 5 - Linear Discriminant Analysis (LDA)"
output:
  html_document:
    df_print: paged
---

# Goal
In this lab, we, again, will try to predict players' skill level, i.e., column `SkillLevel`, which takes integer value between 0 and 3, with 0 being novice level and 3 expert. We will continue to use the `caret` package for this model.


## Brief refresher
`Linear Discriminant Analysis` (LDA) is a method that uses vector space projection to find the best linear decision hyperplanes that separate the classes.

## Preparation

The `caret` package's `LDA` is taken from the `MASS` package.

```{r}
options(repos = c(CRAN = "http://cran.rstudio.com"))
install.packages("MASS", dependencies = TRUE)
install.packages("caret", dependencies = TRUE)
library("MASS")
library("caret")
```

# Step 1: Load the data set 

```{r}
# (optional) set working directory

# read the CSV file
DotaData <- read.csv("DoTalicious_cleaned1000players.csv", fileEncoding="UTF-8-BOM")
summary(DotaData)
```

Clean up and set the correct data type.

```{r}
# convert TotalTime
DotaData$TotalTime <- as.numeric(DotaData$TotalTime)

# exclude the row with SillLevel = SkillLevelNull
DotaData$SkillLevel <- as.factor(DotaData$SkillLevel)
DotaData <- DotaData[!(DotaData$SkillLevel == " SkillLevelNull"),]

# use factor in place of as.factor to remove " SkillLevelNull" from the list of factors
DotaData$SkillLevel <- factor(DotaData$SkillLevel)

# exclude PlayerID
DotaData <- DotaData[-1]

DotaData = na.omit(DotaData)

summary(DotaData)
```

# Step 2: Set up training and test set

This step is again similar to the other labs. 

```{r}
# setting the random seed
set.seed(101)
#Spliting data as training and test set. Using createDataPartition() function from caret
indxTrain <- createDataPartition(y = DotaData$SkillLevel,p = 0.75,list = FALSE)
training <- DotaData[indxTrain,]
testing <- DotaData[-indxTrain,]
summary(training)
summary(testing)
```

## Data Standardization

No need to do this for this particular method.

# Step 3: Apply the LDA model

Now, we are applying the `LDA` model on the data and apply it to predict class values for the `testing` data.

## Train

Use the same `train` function, but change the method to *lda* instead.

```{r}

set.seed(101)
ctrl <- trainControl(method="repeatedcv", number = 10, repeats = 3)
lda_model <- train(SkillLevel ~ ., data = training, method = "lda", trControl = ctrl, tuneLength = 20)
```


```{r}
lda_model
```
No tuning was necessary. The accuracy appears to be low.

## Predict

```{r}
predicted <- predict(lda_model, newdata = testing)
confusionMatrix(predicted, testing$SkillLevel )
```

Results is not as good as with other models we examined before. Performance in predicting `SkillLevel` 1 is particularly weak.


# Conclusion
In this lab, we showed how to use `Linear Discriminant Analysis`, using the *caret* package's wrapper of the `LDA` implementation from the *MASS* package. We showed how `LDA` can be used for prediction. You can compare its results to other models in the previous labs. 

<!-- # Appendix: R Markdown Instructions -->
<!-- Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*. -->

<!-- When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file). -->
